Component response rate variation underlies the stability of highly complex finite systems

### 2020-03-23 notebook/ms.Rmd
Change the title, as requested by Reviewer 2
Insert two additional references that were recommended by reviewer 3

---
title: "Component response rate variation underlies the stability of complex systems"
author: "A. Bradley Duthie    ( alexander.duthie@stir.ac.uk )"
date: Biological and Environmental Sciences, University of Stirling, Stirling, UK,
  FK9 4LA
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
  html_document: default
  word_document:
    fig_caption: yes
    pandoc_args:
    - --csl
    - nature.csl
    reference_docx: docx_template.docx
bibliography: references.bib
header-includes:
- \usepackage{amsmath}
- \usepackage{natbib}
- \usepackage{lineno}
- \usepackage[utf8]{inputenc}
- \linenumbers
- \bibliographystyle{amnatnat}
linestretch: 1
link-citations: yes
linkcolor: blue
csl: nature.csl
biblio-style: apalike
---

@@

In 1972, May [@May1972] first demonstrated that randomly assembled systems of sufficient complexity are almost inevitably unstable given infinitesimally small perturbations. Complexity in this case is defined by the size of the system (i.e., the number of potentially interacting components; $S$), its connectance (i.e., the probability that one component will interact with another; $C$), and the variance of interaction strengths ($\sigma^{2}$) [@Allesina2012]. May's finding that the probability of local stability falls to near zero given a sufficiently high threshold of $\sigma\sqrt{SC}$ is broadly relevant for understanding the dynamics and persistence of systems such as ecological [@May1972; @Allesina2012; @Mougi2012; @Allesina2015; @Grilli2017], neurological [@Gray2008; @Gray2009], biochemical [@Rosenfeld2009; @MacArthur2010], and socio-economic [e.g., banking; @May2008; @Haldane2011; @Suweis2014; @Bardoscia2017] networks. As such, identifying general principles that affect stability in complex systems is of wide-ranging importance.

@@

**Effect of $\mathbf{\sigma^{2}_{\gamma}}$ on $\mathbf{M}$ (co)variation**. The value of $\Re(\lambda_{max})$, and therefore system stability, can be estimated from five properties of $\mathbf{M}$ [@Tang2014b]. These properties include (1) system size ($S$), (2) mean self-regulation of components ($d$), (3) mean interaction strength between components ($\mu$), (4) the variance of between component interaction strengths (hereafter $\sigma^{2}_{M}$, to distinguish from $\sigma^{2}_{A}$ and $\sigma^{2}_{\gamma}$), and (5) the correlation of interaction strengths between components, $M_{ij}$ and $M_{ji}$ ($\rho$). Positive $\sigma^{2}_{\gamma}$ does not change $S$, nor does it necessarily change $E[d]$ or $E[\mu]$. What $\sigma^{2}_{\gamma}$ does change is the total variation in component interaction strengths ($\sigma^{2}_{M}$), and $\rho$. Introducing variation in $\gamma$ increases the total variation in the system. Variation in the off-diagonal elements of $\mathbf{M}$ is described by the joint variation of two random variables,

>>>>

---
title: "Component response rate variation underlies the stability of highly complex finite systems"
author: "A. Bradley Duthie    ( alexander.duthie@stir.ac.uk )"
date: Biological and Environmental Sciences, University of Stirling, Stirling, UK,
  FK9 4LA
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
  html_document: default
  word_document:
    fig_caption: yes
    pandoc_args:
    - --csl
    - nature.csl
    reference_docx: docx_template.docx
bibliography: references.bib
header-includes:
- \usepackage{amsmath}
- \usepackage{natbib}
- \usepackage{lineno}
- \usepackage[utf8]{inputenc}
- \linenumbers
- \bibliographystyle{amnatnat}
linestretch: 1
link-citations: yes
linkcolor: blue
csl: nature.csl
biblio-style: apalike
---

@@

In 1972, May [@May1972] first demonstrated that randomly assembled systems of sufficient complexity are almost inevitably unstable given infinitesimally small perturbations. Complexity in this case is defined by the size of the system (i.e., the number of potentially interacting components; $S$), its connectance (i.e., the probability that one component will interact with another; $C$), and the variance of interaction strengths ($\sigma^{2}$) [@Allesina2012]. May's finding that the probability of local stability falls to near zero given a sufficiently high threshold of $\sigma\sqrt{SC}$ is broadly relevant for understanding the dynamics and persistence of systems such as ecological [@May1972; @Townsend2010a; @Allesina2012; @Mougi2012; @Allesina2015; @Grilli2017], neurological [@Gray2008; @Gray2009], biochemical [@Rosenfeld2009; @MacArthur2010], and socio-economic [e.g., banking; @May2008; @Haldane2011; @Suweis2014; @Bardoscia2017] networks. As such, identifying general principles that affect stability in complex systems is of wide-ranging importance.

@@

**Effect of $\mathbf{\sigma^{2}_{\gamma}}$ on $\mathbf{M}$ (co)variation**. The value of $\Re(\lambda_{max})$, and therefore system stability, can be estimated from five properties of $\mathbf{M}$ [@Tang2014b]. These properties include (1) system size ($S$), (2) mean self-regulation of components ($d$), (3) mean interaction strength between components ($\mu$), (4) the variance of between component interaction strengths (hereafter $\sigma^{2}_{M}$, to distinguish from $\sigma^{2}_{A}$ and $\sigma^{2}_{\gamma}$), and (5) the correlation of interaction strengths between components, $M_{ij}$ and $M_{ji}$ ($\rho$) [@Sommers1988]. Positive $\sigma^{2}_{\gamma}$ does not change $S$, nor does it necessarily change $E[d]$ or $E[\mu]$. What $\sigma^{2}_{\gamma}$ does change is the total variation in component interaction strengths ($\sigma^{2}_{M}$), and $\rho$. Introducing variation in $\gamma$ increases the total variation in the system. Variation in the off-diagonal elements of $\mathbf{M}$ is described by the joint variation of two random variables,
